Left-linear rules in adhesive categories

### 2024-02-10 icalp/fusione.tex
minor things

  If $\der{D}=\{\dder{D}_i\}_{i=0}^n$ and
  $\der{D}'=\{\dder{D}'_i\}_{i=0}^n$ with assciated sequence of rules
  $r(\der{D})=\{\rho_i\}_{i=0}^n$ and
  $r(\der{D}')=\{\rho'_i\}_{i=0}^n$, a \emph{consistent permutation}
  between $(\der{D}, \alpha, \omega)$ and
  $(\der{D}', \alpha', \omega')$ is a permutation
  $\sigma\colon [0,n]\to [0,n]$ such that, for every $i\in [0,n]$,
  $\rho_i=\rho'_{\sigma(i)}$ and, moreover, there exists a
  \emph{mediating isomorphism}
  $\xi_\sigma\colon \tpro{D} \to \lpro \der{D}' \rpro$ fitting in the
  following diagrams, where $m_i, m'_i, h_i$ and $h'_i$ are,
  respectively, the matches and back-matches of $\dder{D}_i$ and
  $\dder{D}'_i$.

@@

  A derivation $\der{D}$ consisting of three steps $\der{D}_0$,
  $\der{D}_1$, and $\der{D}_2$, each applying the corresponding rule
  $\rho_i$ ,is depicted in in Fig.~\ref{fi:derD}. The second and third
  steps $\der{D}_1$ and $\der{D}_2$ are clearly sequential
  independent, via an independence pairs that map nodes consistently
  with their numbering.\todo{P: having names for the graphs would
    probably help}

@@

  The category $\cat{Graph}$ is tame by the result in ...  Thus,
  sequential independence ensure the existence of a switch $\der{E}$,
  depicted in Fig.~\ref{fi:derE}.
  \begin{figure}
    \input{Figure/derE.tex}
    \caption{The derivation $\der{E}$.}
    \label{fi:derE}
  \end{figure}
\end{example}

@@

\begin{definition}[Good independence pair]
  \label{def:filler}
  Let with $\X$ an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear DPO rewriting system. 
  Let also $(i_0, i_1)$ be an independence pair between two direct
  derivations $\dder{D}_0$, $\dder{D}_1$
  as in the solid part of the diagram below
  \[
    \xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} \ar@/^.20cm/[ddrr]|(.3)\hole|(.55)\hole^(.7){u_0} & R_0
      \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0}
      && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1
      \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} \ar@/_.20cm/[ddll]|(.3)\hole|(.55)\hole_(.7){u_1} && R_1 \ar[d]^{h_1} \\
      %
      G_0 &&
      \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1 &&
      \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2\\
      %
      & & & & P \ar@/^.3cm/@{>->}[ull]^{p_0} \ar@/_.3cm/[urr]_{p_1}
    }
  \]

  Consider the pullback of $g_0$ and $f_1$, getting $p_i : P \to D_i$
  ($i \in \{0,1\}$) and the arrows $u_0\colon K_0\to P$ and
  $u_1\colon K_1\to P$ given by \Cref{prop:tec} .
  %
  We will say that $(i_0, i_1)$ is a \emph{good independence pair} if
  the first and second square below are pushouts and if the third
  square exists and it is a pushout too
  \[
    \xymatrix{
      %
      K_0 \ar[r]^{r_0} \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1
      \ar@{>->}[r]^{l_1} \ar[d]_{u_1}& L_1 \ar[d]^{i_1}&K_1
      \ar[r]^{r_1} \ar[d]_{u_1}& R_1\ar@{.>}[d]^{j_0} \\
      %
      P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0} & D_0&P \ar@{.>}[r]_{q_1} &
      Q_1}
  \]
\end{definition}

>>>>

  If $\der{D}=\{\dder{D}_i\}_{i=0}^n$ and
  $\der{D}'=\{\dder{D}'_i\}_{i=0}^n$ with associated sequence of rules
  $r(\der{D})=\{\rho_i\}_{i=0}^n$ and
  $r(\der{D}')=\{\rho'_i\}_{i=0}^n$, a \emph{consistent permutation}
  between $(\der{D}, \alpha, \omega)$ and
  $(\der{D}', \alpha', \omega')$ is a permutation
  $\sigma\colon [0,n]\to [0,n]$ such that, for every $i\in [0,n]$,
  $\rho_i=\rho'_{\sigma(i)}$ and, moreover, there exists a
  \emph{mediating isomorphism}
  $\xi_\sigma\colon \tpro{D} \to \lpro \der{D}' \rpro$ fitting in the
  following diagrams, where $m_i, m'_i, h_i$ and $h'_i$ are,
  respectively, the matches and back-matches of $\dder{D}_i$ and
  $\dder{D}'_i$.

@@

  A derivation $\der{D}$ consisting of three steps $\dder{D}_0$,
  $\dder{D}_1$, and $\dder{D}_2$, each applying the corresponding rule
  $\rho_i$ ,is depicted in in Fig.~\ref{fi:derD}. The second and third
  steps $\dder{D}_1$ and $\dder{D}_2$ are clearly sequential
  independent, via an independence pairs that map nodes consistently
  with their numbering.\todo{P: having names for the graphs would
    probably help}

@@

  The independence pair can be used to construct a switch $\der{E}$, as
  depicted in Fig.~\ref{fi:derE}.
  \begin{figure}
    \input{Figure/derE.tex}
    \caption{The derivation $\der{E}$.}
    \label{fi:derE}
  \end{figure}
\end{example}

@@

\todo{P: la chiamiamo ``strong indepedence pair''? o ``filler pair''}

\begin{definition}[Good independence pair]
  \label{def:filler}
  Let with $\X$ an $\mathcal{M}$-adhesive category 
  and $(\X, \R)$ a left-linear DPO rewriting system. 
  Let also $(i_0, i_1)$ be an independence pair between two direct
  derivations $\dder{D}_0$, $\dder{D}_1$
  as in the solid part of the diagram below
  \[
    \xymatrix@C=18pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} \ar@{->}@/^.20cm/[ddrr]|(.3)\hole|(.55)\hole_(.7){u_0} & R_0
      \ar@/^.25cm/[drrr]|(.3)\hole^(.2){i_0} \ar[dr]|(.3)\hole_{h_0}
      && L_1 \ar@/_.25cm/[dlll]_(.2){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1
      \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} \ar@{->}@/_.20cm/[ddll]|(.3)\hole|(.55)\hole^(.7){u_1} && R_1 \ar[d]^{h_1} \\
      %
      G_0 &&
      \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_(.3){g_0}&& G_1 &&
      \ar@{>->}[ll]^(.3){f_1} D_1 \ar[rr]_{g_1}&& G_2\\
      %
      & & & & P \ar@/^.3cm/@{>-->}[ull]^{p_0} \ar@/_.3cm/@{-->}[urr]_{p_1}
    }
  \]

  Consider the pullback of $g_0$ and $f_1$, getting $p_i : P \to D_i$
  ($i \in \{0,1\}$) and the mediating arrows $u_0\colon K_0\to P$ and
  $u_1\colon K_1\to P$ into the pullback.
  %given by \Cref{prop:tec} .
  %
  We will say that $(i_0, i_1)$ is a \emph{good independence pair} if
  the resulting squares, depicted  
  below, are pushouts and the pushout of $r_1 : K_1 \to R_1$ and $u_1 : K_1 \to P$ exists.
  \[
    \xymatrix@R=5pt{
      %
      K_0 \ar[r]^{r_0} \ar[d]_{u_0}& R_0 \ar[d]^{i_0} & K_1
      \ar@{>->}[r]^{l_1} \ar[d]_{u_1}& L_1 \ar[d]^{i_1}
      % &K_1 \ar[r]^{r_1} \ar[d]_{u_1}& R_1\ar@{.>}[d]^{j_0}
      \\
      %
      P \ar[r]_{p_1} & D_1 &P \ar[r]_{p_0} & D_0
      % & P \ar@{.>}[r]_{q_1} &
      Q_1}
  \]
\end{definition}

The category $\cat{Graph}$ is tame by the result in ..., while
example~\ref{ex:diff1} shows an adhesive non-tame category.
